---
title: "Mixturesof Gaussians"
author: "Antonio, H. F."
date: '2022-11-26'
output: pdf_document
---

Example of a bimodal mixture of Gaussians

```{r}
# Mixture of univariate Gaussians, bimodal
x = seq(-5, 12, length=100)
y = 0.6*dnorm(x, 0, 1) + 0.4*dnorm(x, 5, 2)
par(mar=c(4,4,1,1)+0.1)
plot(x, y, type="l", ylab="Density", las=1, lwd=2)
```



Example of unimodal and skewed mixture of Gaussians

```{r}
# Mixture of univariate Gaussians, unimodal skewed
x = seq(-5, 12, length=100)
y = 0.55*dnorm(x, 0, sqrt(2)) + 0.45*dnorm(x, 3, 4)
par(mar=c(4,4,1,1)+0.1)
plot(x, y, type="l", ylab="Density", las=1, lwd=2)
```

Example of a unimodal, symmetric and heavy tailed mixture Gaussians

```{r}
# Mixture of univariate Gaussians, unimodal heavy tail
x = seq(-12, 12, length=100)
y = 0.40*dnorm(x, 0, sqrt(2)) + 
    0.40*dnorm(x, 0, sqrt(16)) + 
    0.20*dnorm(x, 0, sqrt(20))
z = dnorm(x, 0, sqrt(0.4*2+0.4*16+0.2*20))
par(mar=c(4,4,1,1)+0.1)
plot(x, y, type="l", ylab="Density", las=1, lwd=2)
lines(x, z, lty=2, lwd=2)
legend(2, 0.16, c("Mixture","Gaussian"), lty=c(1,2), bty="n", 
       cex=0.77, lwd=c(2,2))
```

Example of a zero-inflated negative binomial distribution

```{r}
# Zero inflated negative binomial distribution
x = seq(0, 15)
y = dnbinom(x, 8, 0.6)
z = 0.2*c(1,rep(0,length(x)-1)) + (1-0.2)*y
par(mfrow=c(2,1))
par(mar=c(4,4,2,2)+0.1)
barplot(y, names.arg=x, las=1, xlab = "x", ylab="Probability", 
        border=NA, main="Negative Binomial")
par(mar=c(4,4,1,1)+0.1)
barplot(z, names.arg=x, las=1, xlab = "x", ylab="Probability", 
        border=NA, main="Zero-inflated Negative Binomial")

```

Example of zero-inflated log Gaussian distribution

```{r}
#Zero-inflated log Gaussian distribution
x = seq(-2, 15, length=1000)
y = plnorm(x, 1.5, 0.5)
z = 0.3*as.numeric(x>=0) + (1-0.3)*y
par(mar=c(4,4,1,1)+0.1)
plot(x, y, type="l", las=1, lty=2, xlab="x", 
     ylab="Cumulative distribution Function", lwd=2)
lines(x, z, lty=1, lwd=2)
legend(4, 0.45, c("Zero infla. log Gaussian","log Gaussian"), 
     lty=c(1,2), bty="n", lwd=c(2,2))

```



Sample code for simulating from a Mixture

```{r}
# Generate n observations from a mixture of two Gaussian 
# distributions
n     = 50           # Size of the sample to be generated
w     = c(0.6, 0.4)  # Weights
mu    = c(0, 5)      # Means
sigma = c(1, 2)      # Standard deviations
cc    = sample(1:2, n, replace=T, prob=w)
x     = rnorm(n, mu[cc], sigma[cc])
```


```{r}
# Plot f(x) along with the observations 
# just sampled
xx = seq(-5, 12, length=200)
yy = w[1]*dnorm(xx, mu[1], sigma[1]) + 
     w[2]*dnorm(xx, mu[2], sigma[2])
par(mar=c(4,4,1,1)+0.1)
plot(xx, yy, type="l", ylab="Density", xlab="x", las=1, lwd=2)
points(x, y=rep(0,n), pch=1)
points(x, y=rep(0,n), pch=1, col = cc)
```




Advanced simulation and likehood function

Original

```{r}
# Generate n observations from a mixture of two Gaussian 
# distributions
n     = 50           # Size of the sample to be generated
w     = c(0.6, 0.4)  # Weights
mu    = c(0, 5)      # Means
sigma = c(1, 2)      # Standard deviations
cc    = sample(1:2, n, replace=T, prob=w)
x     = rnorm(n, mu[cc], sigma[cc])
    
# Plot f(x) along with the observations 
# just sampled
xx = seq(-5, 12, length=200)
yy = w[1]*dnorm(xx, mu[1], sigma[1]) + 
     w[2]*dnorm(xx, mu[2], sigma[2])
par(mar=c(4,4,1,1)+0.1)
plot(xx, yy, type="l", ylab="Density", xlab="x", las=1, lwd=2)
points(x, y=rep(0,n), pch=1)  
```

Modificated


```{r}
# Generate n observations from a mixture of three Poisso 
# distributions
n     = 200               # Size of the sample to be generated
w     = c(0.7, 0.2, 0.1)  # Weights
mu    = c(1, 2, 6)        # Means
cc    = sample(1:3, n, replace=T, prob=w)
x     = rpois(n, mu[cc])
    
# Plot f(x) along with the observations 
# just sampled
xx = seq.int(0, 20,1)
yy = w[1]*dpois(xx, mu[1]) + 
     w[2]*dpois(xx, mu[2]) + 
     w[3]*dpois(xx, mu[3])
par(mar=c(4,4,1,1)+0.1)
barplot(yy, ylab="Density", xlab="x", las=1, lwd=2)
points(x, y=rep(0,n), pch=1)  

## Option from classmates
empfreq = table(factor(x, levels=seq(0, max(x))))/n
empfreq = table(x)/n
barplot(empfreq)

```



